### 2023-08-11 freechains/rebuttal.md
...

# Rebuttal Letter

We would like to thank the reviewers for the thoughtful comments, questions,
and criticism.

We accommodate all clarifications and answers in the final version of the
paper.

The link that follows shows the *diff* to reach the final version:

- TODO

Each comment is addressed individually as follows.

## Reviewer 1

> Quanto a solução, senti falta de alguma discussão sobre as limitações.
> Em especial, a falta de finalidade, i.e., um post pode ser excluído devido
> aos forks (no momento do merge).
> Este post poderia ser de um usuário legitimo ? quanto isso ocorre e quais as
> implicações?

Sim, para isso, ele teria que escrever em um fork "obscuro", desconectado do
resto dos usuários legítimos.
Dessa forma, ao ser reconectado, esse fork poderia sim ter posts removidos.
A única implicação é que ele teria que postar novamente a mesma mensagem.

No meio da Seção 2.1, essa possibilidade é levantada, mas não com tanta ênfase:

> While applying the branches in order, if any post operation fails, all
> remaining posts are rejected and removed from the DAG, as if they never
> existed.

Em termos práticos, os usuários legítimos se conectam mais diretamente ao
longo do tempo, o que deveria ser suficiente para evitar essa situação.
Para a versão final, nós adicionamos o seguinte trecho:

> Legitimate posts in secondary branches might be removed due to merges, which
> requires that users repost the messages. Nevertheless, we expect that
> legitimate users gossip among themselves more often, thus preventing this
> situation.

> Além disso, é possível que um usuário malicioso participe de mais de uma
> cadeia em um fork? neste caso como fica a regra de merge? e se der empate,
> todos devem escolher o mesmo fork para começar, como isso seria feito?

A reputação de um usuário em uma cadeia X não é válida em uma cadeia Y.
Portanto, interpretamos o comentário acima como "...participe de mais de um
fork em uma mesma cadeia".

Sim, se o usuário malicioso dispor de 1 rep, ele pode escrever propositalmente
em vários forks.
Nesse caso, a reputação prévia dele vai contar igualmente em todos os merges,
se anulando e, portanto, não influenciando no resultado final do merge.

Em caso de empate, que pode acontecer mesmo sem usuários maliciosos, o hash do
fork será critério considerado por todos igualmente.
Esse critério é descrito no final da Seção 2.2:

> ...otherwise, branches use the arbitrary criteria of lexicographical order of
> the post hashes immediately after the common prefix.

Como um possível ataque, o usuário malicioso poderia escrever em vários
branches antes de usuários legítimos, de forma que os merges acabariam por
remover mensagens legítimas.
Essa situação, que complementa o comentário anterior, obrigaria os usuários a
repostarem as mensagens.
Da mesma forma, nós entendemos que os usuários legítimos estarão mais
conectados, inviabilizando esse tipo de ataque.

> Pela descrição, aparentemente na Figura 3 deveria ter uma transição de
> revoked  para accepted.

A Figura 3 já tem essa transição.
Há uma seta bi-direcional entre os dois estados.
(Espero ter entendido corretamente esse comentário.)

# Reviewer 2

> It is strange that known techniques like EigenTrust are not even mentioned,
> even though we are in the blockchain domain.

TODO

> We do not know if the results are good or bad because there was no comparison
> to the state of the art or baseline.

The main goal of our experiments in Section 3.1 is to stress the consensus
algorithm to show that the mechanism is feasible, i.e., that it has a
reasonable performance to be used in practice.
The good result is that the protocol works in a permissionless context,
regardless of the slower performance.

We added the following remark at the beginning of Section 3.1:

> The main goal of this section is to stress the consensus algorithm to show
> that permissionless public forums are viable, regardless of the inherent
> slower performance in comparison to permissioned protocols.

> Also, at the end of the day, we do not know if the proposal is resistant to
> Sybil attacks because there was no validation of the consensus and reputation
> mechanisms.

In our context, a Sybil is a random machine/identity with no previous
reputation in the chains, which, by definition, cannot operate in the chains.
By default, users with no reps do not have write access, as described in the
beginning of Section 2:

> To prevent Sybils, users with no *reps* cannot operate under these rules,
> requiring a welcoming like from any other user already in the system.
> The fact that likes are zero-sum operations, which only transfer reputation,
> eliminates the incentives from malicious users to invite Sybils into the
> system.
> The only way to generate new *reps* is to post content that other users
> tolerate, which demands non-trivial work resistant to automation.

We included the following remarks:

> We consider Sybils to be groups of throw-away identities and machines with
> no previous reputation in the forums.
> ...
> In this sense, having one or hundreds of machines in the network does not
> affect the capacity to write into the forums.
> ...
> In our analogy with Bitcoin, Sybil attacks would require to add human
> resources, instead of CPU power.

# Reviewer 3

> It's important to note that the reputation system does not enforce "good"
> human behavior but provides a transparent and quantitative way for users to
> understand forum evolution and act accordingly. 
> ...
> I am also concerned that the content removal is too weak to work in practice.
> For example, a Forum may be taken over by extremists, and they can upvote
> illegal content. That would require removal due to legal process, or face
> blocking at network level.

This is all true, and is the reason why we close the paper with your remark
above.
It is not even required to be "taken by extremists", but only "created by and
for extremists".
But we also believe that a protocol that is tamper proof, and provides
persistence and data transparency can actually support law enforcement.

> Maybe a way to go is to create a pool of super-reputations that could answer
> to injunctions and administratively remove the content and a measure to avoid
> illegal content being broadcasted in subverted grupus.

For extreme conflicts, in which content removal is too weak, the protocol also
provides hard forks, as discussed in the middle of Section 2.2:

> More than simple numeric disputes, a hard fork represents a social conflict
> in which reconciling branches is no longer possible.

> The paper is too descriptive, so reading sometime is tiresome.
> ...
> Reading could be improves by using some figures to explain ideas. Sometimes
> it is difficult to follow the text.

We opted for a full description of the protocol, but we recognize that the
paper became a bit dense.
It is hard to modify the whole text without more specific suggestions, but we
removed some details of Freechains from Section 3.